Automatic Matching of Bullet Lands

### 2016-07-11 imaging-paper.Rnw
Some small typos

ggplot() + theme_bw() +
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
            data=peaks,  alpha=0.2) +
  geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks, alpha=.7) +
  scale_colour_brewer(palette="Set2") + 
  scale_fill_brewer(palette="Set2") +
  theme(legend.position="none") + 
  facet_grid(smoothfactor~., labeller="label_both") +
  geom_line(aes(x=y, y=resid, group=x), data=br111.loess$data,  
            colour="grey50") +
  geom_line(aes(x=y, y=smoothed), data=smooths) +
  ylab(expression(paste("Signature values (in ",mu,"m)", sep=""))) 
@
\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 25, and 45, respectively. Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 45 might over-smooth and cause the peaks/valleys to either end disappear or shift  horizontally  from their position in the signature.}
\end{figure}

@@

Using recursive partitioning, we fit a decision tree~\citep{breiman:1984, rpart, rpart.plot} to predict matches between lands based on features derived from the image files. The resulting tree is shown in Figure~\ref{fig:tree}. A total of \Sexpr{sum(bstats$match[bstats$pred>0.5])} lands is being matched correctly. Interestingly, the number of consecutive matching striae does not feature in this evaluation. 
Instead of CMS, cross-correlation (ccf) between the signatures is very important in the matching process  by the decision tree. Aside from  cross correlation, the total number of matches is also included in the decision rule. 
Between cross-correlation and CMS, cross-correlation has higher  predictive power. This  does not  contradict earlier findings emphasizing the value of CMS on visual assessments of bullet matches: in those papers, assessments were based on purely visual inspection of either actual bullets or 2D microscopic images of bullets.
Neither one of these methods allows for an assessment of cross-correlations. This is one of the benefits of switching to a digitized version of the images that preserves the 3D surface structure. The findings about the discriminating power of cross-correlation is consistent with the results of the study by \citet{ma:2004}. However, in that study, authors did not consider the number of matches and non-matches.

@@

However, error rates based on bullet-to-bullet matches do not carry a lot of weight because of the small size of the study: fifteen unknown bullets are successfully matched to two pairs of ten bullets. Matching bullets can only be tested realistically in a much bigger experiment. 
Another thing to note about  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.\ whenever the predicted probability of a match exceeds 0.5, a match is declared. Basing this decision on 0.5 is not necessary. In practice, examiners are allowed a third option of `inconclusive'. On a probability spectrum of outcomes we could therefore introduce an interval of `inconclusive' results in the middle of the spectrum -- which turns out to be not necessary in the Hamby study, because, here, the results from the random forest are very clear cut. Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree and the random forest. %\hh{The tree predicts 132 known matches correctly, with 40 false negatives and one false positive match. The random forest predicts all 172 known matches correctly as well as all 10,212 known non-matches.} 
%Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only four cases result in a different decision.

>>>>

ggplot() + theme_bw() +
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
            data=peaks,  alpha=0.2) +
  geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks, alpha=.7) +
  scale_colour_brewer(palette="Set2") + 
  scale_fill_brewer(palette="Set2") +
  theme(legend.position="none") + 
  facet_grid(smoothfactor~., labeller="label_both") +
  geom_line(aes(x=y, y=resid, group=x), data=br111.loess$data,  
            colour="grey50") +
  geom_line(aes(x=y, y=smoothed), data=smooths) +
  ylab(expression(paste("Signature values (in ",mu,"m)", sep=""))) 
@
\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 25, and 45, respectively. Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 45 might over-smooth and cause the peaks/valleys to either end disappear or shift  horizontally from their original position in the signature.}
\end{figure}

@@

Using recursive partitioning, we fit a decision tree~\citep{breiman:1984, rpart, rpart.plot} to predict matches between lands based on features derived from the image files. The resulting tree is shown in Figure~\ref{fig:tree}. A total of \Sexpr{sum(bstats$match[bstats$pred>0.5])} lands is being matched correctly. Interestingly, the number of consecutive matching striae does not feature in this evaluation. 
Instead of CMS, cross-correlation (ccf) between the signatures is very important in the matching process  by the decision tree. Aside from  cross correlation, the total number of matches is also included in the decision rule. 
Between cross-correlation and CMS, cross-correlation has higher  predictive power. This  does not  contradict earlier findings emphasizing the value of CMS on visual assessments of bullet matches: in those papers, assessments were based on purely visual inspection of either actual bullets or 2D microscopic images of bullets.
Neither one of these methods allows for an assessment of cross-correlations. This is one of the benefits of switching to a digitized version of the images that preserves the 3D surface structure. The findings about the discriminating power of cross-correlation are consistent with the results of the study by \citet{ma:2004}. However, in that study, the authors did not consider the number of matches and non-matches.

@@

However, error rates based on bullet-to-bullet matches do not carry a lot of weight because of the small size of the study: fifteen unknown bullets are successfully matched to two pairs of ten bullets. Matching bullets can only be tested realistically in a much bigger experiment. 
Another thing to note about  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.\ whenever the predicted probability of a match exceeds 0.5, a match is declared. Basing this decision on 0.5 is not necessary. In practice, examiners are allowed a third option of `inconclusive'. On a probability spectrum of outcomes we could therefore introduce an interval of `inconclusive' results in the middle of the spectrum -- which turns out to be unnecessary in the Hamby study, because, here, the results from the random forest are very clear cut. Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree and the random forest. %\hh{The tree predicts 132 known matches correctly, with 40 false negatives and one false positive match. The random forest predicts all 172 known matches correctly as well as all 10,212 known non-matches.} 
%Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only four cases result in a different decision.
